# Vectors

```{r}
library(tidyverse)
```


## Attributes

Set an attribute with `attr()`.

```{r}
x <- c(1,3,5)
```

```{r}
str(x)
```

```{r}
attr(x, "y") <- "hello"
```

```{r}
x
```

```{r}
str(x)
```

Create an object with attributes using `structure()`

```{r}
my_struc <-
  structure(c(1,3,5),
            color="red",
            name = "max")
```

```{r}
my_struc
```

If you give the `dim` attribute to a vector, it will become a matrix.

```{r}
a <- 1:10

print(a)
```

```{r}
# Give it 2 rows and 5 columns
dim(a) <- c(2,5)
```

```{r}
print(a)
```

A cool fact is that you can create "list-matrices" by passing the `dim` attribute to a list.

```{r}
l <- list(c(1L, 2L, 3L),
          "a",
          TRUE,
          1.0)

print(l)
```

```{r}
dim(l) <- c(2, 2)

print(l)
```

## Factors

You can make a factor by specifying the proper attributes: `class` and `levels`. Note that this will only work on a vector of integers.

```{r}
x <- c(1L, 1L, 2L, 2L, 3L, 3L)

# This won't work because these aren't integers.
# x <- c(1,1,2,2,3,3)
```

```{r}
attributes(x) <- list(class="factor",
                      levels=c("a", "b", "c"))
```

```{r}
# Now we have a factor!
print(x)
```

## Data Frames

Important point:

> A data frame is a named list of vectors with attributes for (column) `names`, `row.names`, and its class, `data.frame`.

> [T]he length of each of its vectors must be the same

`names()` is the same as `colnames()` for data frames.

```{r}
names(mtcars) == colnames(mtcars)
```

The length of a data frame is the number of columns.

```{r}
length(mtcars) == ncol(mtcars)

length(mtcars) != nrow(mtcars)
```

> [T]ibbles will only recycle vectors of length one

> `tibble()` allows you to refer to variables created during construction

Data frames can have list-columns.

```{r}
tibble(x = 1:3,
       y = list(1:2, 1:3, 1:4))
```

Elements of a data frame can also be matrices, arrays, or other data frames!

## NULL

Two uses of `NULL`:

> To represent an empty vector (a vector of length zero) of arbitrary type. For example, if you use `c()` but don’t include any arguments, you get `NULL`, and concatenating `NULL` to a vector will leave it unchanged.

> To represent an absent vector. For example, `NULL` is often used as a default function argument, when the argument is optional but the default value requires some computation. Contrast this with NA which is used to indicate that an element of a vector is absent.




